---
title: "Lab 6"
author: "Lory Salazar"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(ggpubr)
library(broom)
```

## Looking at our sample data
First, we're making mock unpaired data by creating two sample vectors using the set.seed() function.

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

# set.seed is to create a "pseudorandom" sample

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Is there evidence for significant difference in ranks between the populations that these two groups were drawn from?

```{r}
hist(gp_1)

hist(gp_2)
```

The histograms show a not clearly normally distribution, and the sample size is small. So, using a median-based test is probably good to compare these data. 


## Mann-Whitney U Test
We are performing a Mann-Whitney U test to see if there is significant difference between the group medians.

```{r}
my_mwu <- wilcox.test(gp_1, gp_2)
```

The p-value is 0.28, which means that if the null hypothesis is true, there is a probability of 28% that we could have found median values at least this different by chance. 
This is not sufficient evidence to reject the null hypothesis (samples were drawn from populations with the same median) if we use the significance level = 0.05.

Note to use ?kruskal.test for info about a rank-based test for comparing medians across >2 groups.

## Simple linear regression
First, we make an exploratory scatterplot to compare penguin flipper length vs body mass.

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

Looking at this data, we can see there is a linear relationship, so it makes sense to move forward with our test.

### Model it
Model the data using lm().

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

summary(penguin_lm)
```

The R^2 value is 0.759 so 75.9% of variance in body mass is explained by flipper length.

Next, we can use broom::tidy() to get model outputs in nice data frame format. 

```{r}
penguin_lm_tidy <- broom::tidy(penguin_lm)

penguin_int <- penguin_lm_tidy$estimate[1] # this is to get just the intercept

penguin_coef <- penguin_lm_tidy$estimate[2]

penguin_lm_out <- broom::glance(penguin_lm) # this is to look at the stats more easily 
penguin_lm_out
```

Simple linear regression was used to explore the relationship between penguin flipper length (mm) and body mass (g) across all three penguin species, and including both male and female penguins. A significant regression model was found ($\beta$ = `r round(penguin_coef,3)`, F(`r penguin_lm_out$df`,`r penguin_lm_out$df.residual`) = `r round(penguin_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(penguin_lm_out$r.squared,3)`.

## Explore model assumptions

```{r}
plot(penguin_lm) # base R command (?) to get a bunch of plots
```

Interpreting these plots...

- **The first one**: fitted values vs. residuals 
- **The second one**: QQ-plot for residuals 
- **The third one**: another way of looking at fitted vs. residuals (these are just standardized residuals, but you can interpret it the same way)
- **The fourth one**: Cook's distance, a measure of "influence" or "leverage" that individual points have on the model - often considered a way to explore outliers.

## Visualize the model 
Now that we have explored the assumptions and decided liner regression is a valid tool to use for this data, let's look at the model.

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) + # geom_smooth is to add a linear model to a scatterplot
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700) # adds the equation to the plot and labels it
```

## Find Person's r for correlation

```{r}
penguins_cor <- cor.test(penguins$flipper_length_mm, penguins$body_mass_g) # will report the r value which tells us if flipper length and body mass are correlated
```
The r value is 0.87. There is a strong positive correlation between the penguin flipper length and body mass. 
